---
title: "Ebusamba School-Upper Primary (8-4-4) Exam Analysis"
author: Bugame Gift
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    code_folding: show
    theme:
      bg: "#660000"
      fg: "#FFF5EE"
      primary: "#E9DCC9"
      secondary: "#FFF5EE"
      base_font:
        google: Prompt
      heading_font:
        google: Proza Libre
---

```{r setup-chunk, include=FALSE}
library(knitr)
opts_chunk$set(message=FALSE, warning=FALSE, tidy.opts = list(width.cutoff=60))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(pander))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(shinydashboard))


# GGPLOT THEME SET
theme_set(theme_classic())

eb6 = data.frame(# class 6
  read_csv("eb6.csv"), stringsAsFactors = TRUE)

eb7 = data.frame(# class 7
  read_csv("eb7.csv"), stringsAsFactors = TRUE)

eb8 = data.frame(# class 8
  read_csv("eb8.csv"), stringsAsFactors = TRUE)

```

# Introduction {.tabset .tabset-pill}

This report presents the analysis of the *Solution 005* examination done by the upper primary school pupils in standards 6 to 8 in Ebusamba School in 2021. The pupils were examined on 5 subjects and subject was marked out of a possible 100 marks.

# Objectives

-   To describe the subject-wise performance per class
-   To investigate for differences in stream performance per class

------------------------------------------------------------------------

# Subject-wise performance {.tabset .tabset-pill}

The analysis begins with an overview of how the students from each class performed in the 5 subjects.

## Standard 6 {.tabset .tabset-pill}

### a. English {.tabset .tabset-pill}

This is how Standard 6 pupils performed in English.

#### Descriptive summaries

```{r,english6-1, echo=FALSE,fig.align='center'}
descriptive.eng6 = data.frame(
  "Minimum"=round(min(eb6$eng),2),
  "Q1"=round(quantile(eb6$eng, 0.25),2),
  "Mean"=round(mean(eb6$eng),2),
  "Median"=round(median(eb6$eng),2),
  "Q3"=round(quantile(eb6$eng, 0.75),2),
  "Maximum"=round(max(eb6$eng),2),
  "Std.deviation"=round(sd(eb6$eng),2)
  )

descriptive.eng6 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,english6-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.eng6 = cut(eb6$eng,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.eng6)*100/nrow(eb6)) %>% pander()
  
```

#### Density function

```{r,english6-density, echo=FALSE,fig.align='center'}

ggplot(eb6, aes(x=eng))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.5)+
  labs(x="score",
       y="probability",
       title="Standard 6",
       subtitle="English:")+
  geom_vline(xintercept = mean(eb6$eng))

```

#### Best student(s)

```{r,english6-best, echo=FALSE,fig.align='center'}
eb6[eb6$eng == max(eb6$eng),] %>% pander()
```

### b. Kiswahili {.tabset .tabset-pill}

This is how Standard 6 pupils performed in Kiswahili.

#### Descriptive summaries

```{r,kiswahili6-1, echo=FALSE,fig.align='center'}
descriptive.kis6 = data.frame(
  "Minimum"=round(min(eb6$kis),2),
  "Q1"=round(quantile(eb6$kis, probs = 0.25),2),
  "Mean"=round(mean(eb6$kis),2),
  "Median"=round(median(eb6$kis),2),
  "Q3"=round(quantile(eb6$kis, probs = 0.75),2),
  "Maximum"=round(max(eb6$kis),2),
  "Std.deviation"=round(sd(eb6$kis),2)
  )

descriptive.kis6 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,kiswahili6-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.kis6 = cut(eb6$kis,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.kis6)*100/nrow(eb6)) %>% pander()
  
```

#### Density function

```{r,kiswahili6-density, echo=FALSE,fig.align='center'}

ggplot(eb6, aes(x=kis))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.5)+
  labs(x="score",
       y="probability",
       title="Standard 6",
       subtitle="Kiswahili:")+
  geom_vline(xintercept = mean(eb6$kis))
```

#### Best student(s)

```{r,kiswahili6-best, echo=FALSE,fig.align='center'}
eb6[eb6$kis == max(eb6$kis),] %>% pander()
```

### c. Mathematics {.tabset .tabset-pill}

This is how Standard 6 pupils performed in Mathematics.

#### Descriptive summaries

```{r,mathematics6-1, echo=FALSE,fig.align='center'}
descriptive.math6 = data.frame(
  "Minimum"=round(min(eb6$math),2),
  "Q1"=round(quantile(eb6$math, probs = 0.25),2),
  "Mean"=round(mean(eb6$math),2),
  "Median"=round(median(eb6$math),2),
  "Q3"=round(quantile(eb6$math, probs = 0.75),2),
  "Maximum"=round(max(eb6$math),2),
  "Std.deviation"=round(sd(eb6$math),2)
  )

descriptive.math6 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,mathematics6-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.math6 = cut(eb6$math,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.math6)*100/nrow(eb6)) %>% pander()
  
```

#### Density function

```{r,math6-density, echo=FALSE,fig.align='center'}

ggplot(eb6, aes(x=math))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.5)+
  labs(x="score",
       y="probability",
       title="Standard 6",
       subtitle="Mathematics:")+
  geom_vline(xintercept = mean(eb6$math))

```

#### Best student

```{r,math6-best, echo=FALSE,fig.align='center'}
eb6[eb6$math == max(eb6$math),] %>% pander()
```

### d. Science {.tabset .tabset-pill}

This is how Standard 6 pupils performed in Science.

#### Descriptive summaries

```{r,science6-1, echo=FALSE,fig.align='center'}
descriptive.sci6 = data.frame(
  "Minimum"=round(min(eb6$sci),2),
  "Q1"=round(quantile(eb6$sci, probs = 0.25),2),
  "Mean"=round(mean(eb6$sci),2),
  "Median"=round(median(eb6$sci),2),
  "Q3"=round(quantile(eb6$sci, probs = 0.75),2),
  "Maximum"=round(max(eb6$sci),2),
  "Std.deviation"=round(sd(eb6$sci),2)
  )

descriptive.sci6 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,science6-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.sci6 = cut(eb6$sci,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.sci6)*100/nrow(eb6)) %>% pander()
  
```

#### Density function

```{r,science6-density, echo=FALSE,fig.align='center'}

ggplot(eb6, aes(x=sci))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.5)+
  labs(x="score",
       y="probability",
       title="Standard 6",
       subtitle="Science:")+
  geom_vline(xintercept = mean(eb6$sci))
```

#### Best student

```{r,science6-best, echo=FALSE,fig.align='center'}
eb6[eb6$sci == max(eb6$sci),] %>% pander()
```

### e. SSRE {.tabset .tabset-pill}

This is how Standard 6 pupils performed in SSRE.

#### Descriptive summaries

```{r,ssre6-1, echo=FALSE,fig.align='center'}
descriptive.ssre6 = data.frame(
  "Minimum"=round(min(eb6$ssre),2),
  "Q1"=round(quantile(eb6$ssre, probs = 0.25),2),
  "Mean"=round(mean(eb6$ssre),2),
  "Median"=round(median(eb6$ssre),2),
  "Q3"=round(quantile(eb6$ssre, probs = 0.75),2),
  "Maximum"=round(max(eb6$ssre),2),
  "Std.deviation"=round(sd(eb6$ssre),2)
  )

descriptive.ssre6 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,ssre6-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.ssre6 = cut(eb6$ssre,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.ssre6)*100/nrow(eb6)) %>% pander()
  
```

#### Density function

```{r,SSRE6-density, echo=FALSE,fig.align='center'}

ggplot(eb6, aes(x=ssre))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.5)+
  labs(x="score",
       y="probability",
       title="Standard 6",
       subtitle="SSRE:")+
  geom_vline(xintercept = mean(eb6$ssre))
```

#### Best student

```{r,SSRE6-best, echo=FALSE,fig.align='center'}
eb6[eb6$ssre == max(eb6$ssre),] %>% pander()
```

### f. Total {.tabset .tabset-pill}

This is how Standard 6 pupils performed overally.

#### Descriptive summaries

```{r,total6-1, echo=FALSE,fig.align='center'}
descriptive.total6 = data.frame(
  "Minimum"=round(min(eb6$total),2),
  "Q1"=round(quantile(eb6$total, probs = 0.25),2),
  "Mean"=round(mean(eb6$total),2),
  "Median"=round(median(eb6$total),2),
  "Q3"=round(quantile(eb6$total, probs = 0.75),2),
  "Maximum"=round(max(eb6$total),2),
  "Std.deviation"=round(sd(eb6$total),2)
  )

descriptive.total6 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,total6-2, echo=FALSE, fig.align='center'}
total.custom.breaks <-c(0,100,200,300,400,500) 
total.custom.labels <-c("0-100","101-200","201-300","301-400","401-500")

cut.total6 = cut(eb6$total,
                 breaks=total.custom.breaks, 
                 labels=total.custom.labels)

ceiling(table(cut.total6)*100/nrow(eb6)) %>% pander()
  
```

#### Density function

```{r,total6-density, echo=FALSE,fig.align='center'}

ggplot(eb6, aes(x=total))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.5)+
  labs(x="score",
       y="probability",
       title="Standard 6",
       subtitle="Total:")+
  geom_vline(xintercept = mean(eb6$total))

```

#### Best student

```{r,total6-best, echo=FALSE,fig.align='center'}
eb6[eb6$total == max(eb6$total),] %>% pander()
```

## Standard 7 {.tabset .tabset-pill}

### a. English {.tabset .tabset-pill}

This is how Standard 7 pupils performed in English.

#### Descriptive summaries

```{r,english7-1, echo=FALSE,fig.align='center'}
descriptive.eng7 = data.frame(
  "Minimum"=round(min(eb7$eng),2),
  "Q1"=round(quantile(eb7$eng, 0.25),2),
  "Mean"=round(mean(eb7$eng),2),
  "Median"=round(median(eb7$eng),2),
  "Q3"=round(quantile(eb7$eng, 0.75),2),
  "Maximum"=round(max(eb7$eng),2),
  "Std.deviation"=round(sd(eb7$eng),2)
  )

descriptive.eng7 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,english7-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.eng7 = cut(eb7$eng,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.eng7)*100/nrow(eb7)) %>% pander()
  
```

#### Density function

```{r,english7-density, echo=FALSE,fig.align='center'}

ggplot(eb7, aes(x=eng))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.7)+
  labs(x="score",
       y="probability",
       title="Standard 7",
       subtitle="English:")+
  geom_vline(xintercept = mean(eb7$eng))

```

#### Best student(s)

```{r,english7-best, echo=FALSE,fig.align='center'}
eb7[eb7$eng == max(eb7$eng),] %>% pander()
```

### b. Kiswahili {.tabset .tabset-pill}

This is how Standard 7 pupils performed in Kiswahili.

#### Descriptive summaries

```{r,kiswahili7-1, echo=FALSE,fig.align='center'}
descriptive.kis7 = data.frame(
  "Minimum"=round(min(eb7$kis),2),
  "Q1"=round(quantile(eb7$kis, probs = 0.25),2),
  "Mean"=round(mean(eb7$kis),2),
  "Median"=round(median(eb7$kis),2),
  "Q3"=round(quantile(eb7$kis, probs = 0.75),2),
  "Maximum"=round(max(eb7$kis),2),
  "Std.deviation"=round(sd(eb7$kis),2)
  )

descriptive.kis7 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,kiswahili7-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.kis7 = cut(eb7$kis,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.kis7)*100/nrow(eb7)) %>% pander()
  
```

#### Density function

```{r,kiswahili7-density, echo=FALSE,fig.align='center'}

ggplot(eb7, aes(x=kis))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.7)+
  labs(x="score",
       y="probability",
       title="Standard 7",
       subtitle="Kiswahili:")+
  geom_vline(xintercept = mean(eb7$kis))

```

#### Best student(s)

```{r,kiswahili7-best, echo=FALSE,fig.align='center'}
eb7[eb7$kis == max(eb7$kis),] %>% pander()
```

### c. Mathematics {.tabset .tabset-pill}

This is how Standard 7 pupils performed in Mathematics.

#### Descriptive summaries

```{r,mathematics7-1, echo=FALSE,fig.align='center'}
descriptive.math7 = data.frame(
  "Minimum"=round(min(eb7$math),2),
  "Q1"=round(quantile(eb7$math, probs = 0.25),2),
  "Mean"=round(mean(eb7$math),2),
  "Median"=round(median(eb7$math),2),
  "Q3"=round(quantile(eb7$math, probs = 0.75),2),
  "Maximum"=round(max(eb7$math),2),
  "Std.deviation"=round(sd(eb7$math),2)
  )

descriptive.math7 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,mathematics7-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.math7 = cut(eb7$math,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.math7)*100/nrow(eb7)) %>% pander()
  
```

#### Density function

```{r,math7-density, echo=FALSE,fig.align='center'}

ggplot(eb7, aes(x=math))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.7)+
  labs(x="score",
       y="probability",
       title="Standard 7",
       subtitle="Mathematics:")+
  geom_vline(xintercept = mean(eb7$math))

```

#### Best student(s)

```{r,math7-best, echo=FALSE,fig.align='center'}
eb7[eb7$math == max(eb7$math),] %>% pander()
```

### d. Science {.tabset .tabset-pill}

This is how Standard 7 pupils performed in Science.

#### Descriptive summaries

```{r,science7-1, echo=FALSE,fig.align='center'}
descriptive.sci7 = data.frame(
  "Minimum"=round(min(eb7$sci),2),
  "Q1"=round(quantile(eb7$sci, probs = 0.25),2),
  "Mean"=round(mean(eb7$sci),2),
  "Median"=round(median(eb7$sci),2),
  "Q3"=round(quantile(eb7$sci, probs = 0.75),2),
  "Maximum"=round(max(eb7$sci),2),
  "Std.deviation"=round(sd(eb7$sci),2)
  )

descriptive.sci7 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,science7-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.sci7 = cut(eb7$sci,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.sci7)*100/nrow(eb7)) %>% pander()
  
```

#### Density function

```{r,science7-density, echo=FALSE,fig.align='center'}

ggplot(eb7, aes(x=sci))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.7)+
  labs(x="score",
       y="probability",
       title="Standard 7",
       subtitle="Science:")+
  geom_vline(xintercept = mean(eb7$sci))

```

#### Best student(s)

```{r,science7-best, echo=FALSE,fig.align='center'}
eb7[eb7$sci == max(eb7$sci),] %>% pander()
```

### e. SSRE {.tabset .tabset-pill}

This is how Standard 7 pupils performed in SSRE.

#### Descriptive summaries

```{r,ssre7-1, echo=FALSE,fig.align='center'}
descriptive.ssre7 = data.frame(
  "Minimum"=round(min(eb7$ssre),2),
  "Q1"=round(quantile(eb7$ssre, probs = 0.25),2),
  "Mean"=round(mean(eb7$ssre),2),
  "Median"=round(median(eb7$ssre),2),
  "Q3"=round(quantile(eb7$ssre, probs = 0.75),2),
  "Maximum"=round(max(eb7$ssre),2),
  "Std.deviation"=round(sd(eb7$ssre),2)
  )

descriptive.ssre7 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,ssre7-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.ssre7 = cut(eb7$ssre,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.ssre7)*100/nrow(eb7)) %>% pander()
  
```

#### Density function

```{r,ssre7-density, echo=FALSE,fig.align='center'}

ggplot(eb7, aes(x=ssre))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.7)+
  labs(x="score",
       y="probability",
       title="Standard 7",
       subtitle="SSRE:")+
  geom_vline(xintercept = mean(eb7$ssre))

```

#### Best student(s)

```{r,ssre7-best, echo=FALSE,fig.align='center'}
eb7[eb7$ssre == max(eb7$ssre),] %>% pander()
```

### f. Total {.tabset .tabset-pill}

This is how Standard 7 pupils performed overally.

#### Descriptive summaries

```{r,total7-1, echo=FALSE,fig.align='center'}
descriptive.total7 = data.frame(
  "Minimum"=round(min(eb7$total),2),
  "Q1"=round(quantile(eb7$total, probs = 0.25),2),
  "Mean"=round(mean(eb7$total),2),
  "Median"=round(median(eb7$total),2),
  "Q3"=round(quantile(eb7$total, probs = 0.75),2),
  "Maximum"=round(max(eb7$total),2),
  "Std.deviation"=round(sd(eb7$total),2)
  )

descriptive.total7 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,total7-2, echo=FALSE, fig.align='center'}
total.custom.breaks <-c(0,100,200,300,400,500) 
total.custom.labels <-c("0-100","101-200","201-300","301-400","401-500")

cut.total7 = cut(eb7$total,
                 breaks=total.custom.breaks, 
                 labels=total.custom.labels)

ceiling(table(cut.total7)*100/nrow(eb7)) %>% pander()
  
```

#### Density function

```{r,total7-density, echo=FALSE,fig.align='center'}

ggplot(eb7, aes(x=total))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.7)+
  labs(x="score",
       y="probability",
       title="Standard 7",
       subtitle="Total:")+
  geom_vline(xintercept = mean(eb7$total))

```

#### Best student(s)

```{r,total7-best, echo=FALSE,fig.align='center'}
eb7[eb7$total == max(eb7$total),] %>% pander()
```

## Standard 8 {.tabset .tabset-pill}

### a. English {.tabset .tabset-pill}

This is how Standard 8 pupils performed in English.

#### Descriptive summaries

```{r,english8-1, echo=FALSE,fig.align='center'}
descriptive.eng8 = data.frame(
  "Minimum"=round(min(eb8$eng),2),
  "Q1"=round(quantile(eb8$eng, 0.25),2),
  "Mean"=round(mean(eb8$eng),2),
  "Median"=round(median(eb8$eng),2),
  "Q3"=round(quantile(eb8$eng, 0.75),2),
  "Maximum"=round(max(eb8$eng),2),
  "Std.deviation"=round(sd(eb8$eng),2)
  )

descriptive.eng8 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,english8-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.eng8 = cut(eb8$eng,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.eng8)*100/nrow(eb8)) %>% pander()
  
```

#### Density function

```{r,english8-density, echo=FALSE,fig.align='center'}

ggplot(eb8, aes(x=eng))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.9)+
  labs(x="score",
       y="probability",
       title="Standard 8",
       subtitle="English:")+
  geom_vline(xintercept = mean(eb8$eng))

```

#### Best student(s)

```{r,eng8-best, echo=FALSE,fig.align='center'}
eb8[eb8$eng == max(eb8$eng),] %>% pander()
```

### b. Kiswahili {.tabset .tabset-pill}

This is how Standard 8 pupils performed in Kiswahili.

#### Descriptive summaries

```{r,kiswahili8-1, echo=FALSE,fig.align='center'}
descriptive.kis8 = data.frame(
  "Minimum"=round(min(eb8$kis),2),
  "Q1"=round(quantile(eb8$kis, probs = 0.25),2),
  "Mean"=round(mean(eb8$kis),2),
  "Median"=round(median(eb8$kis),2),
  "Q3"=round(quantile(eb8$kis, probs = 0.75),2),
  "Maximum"=round(max(eb8$kis),2),
  "Std.deviation"=round(sd(eb8$kis),2)
  )

descriptive.kis8 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,kiswahili8-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.kis8 = cut(eb8$kis,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.kis8)*100/nrow(eb8)) %>% pander()
  
```

#### Density function

```{r,kiswahili8-density, echo=FALSE,fig.align='center'}

ggplot(eb8, aes(x=kis))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.9)+
  labs(x="score",
       y="probability",
       title="Standard 8",
       subtitle="Kiswahili:")+
  geom_vline(xintercept = mean(eb8$kis))

```

#### Best student(s)

```{r,kiswahili8-best, echo=FALSE,fig.align='center'}
eb8[eb8$kis == max(eb8$kis),] %>% pander()
```

### c. Mathematics {.tabset .tabset-pill}

This is how Standard 8 pupils performed in Mathematics.

#### Descriptive summaries

```{r,mathematics8-1, echo=FALSE,fig.align='center'}
descriptive.math8 = data.frame(
  "Minimum"=round(min(eb8$math),2),
  "Q1"=round(quantile(eb8$math, probs = 0.25),2),
  "Mean"=round(mean(eb8$math),2),
  "Median"=round(median(eb8$math),2),
  "Q3"=round(quantile(eb8$math, probs = 0.75),2),
  "Maximum"=round(max(eb8$math),2),
  "Std.deviation"=round(sd(eb8$math),2)
  )

descriptive.math8 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,mathematics8-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.math8 = cut(eb8$math,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.math8)*100/nrow(eb8)) %>% pander()
  
```

#### Density function

```{r,math8-density, echo=FALSE,fig.align='center'}

ggplot(eb8, aes(x=math))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.9)+
  labs(x="score",
       y="probability",
       title="Standard 8",
       subtitle="Mathematics:")+
  geom_vline(xintercept = mean(eb8$math))

```

#### Best student(s)

```{r,math8-best, echo=FALSE,fig.align='center'}
eb8[eb8$math == max(eb8$math),] %>% pander()
```

### d. Science {.tabset .tabset-pill}

This is how Standard 8 pupils performed in Science.

#### Descriptive summaries

```{r,science8-1, echo=FALSE,fig.align='center'}
descriptive.sci8 = data.frame(
  "Minimum"=round(min(eb8$sci),2),
  "Q1"=round(quantile(eb8$sci, probs = 0.25),2),
  "Mean"=round(mean(eb8$sci),2),
  "Median"=round(median(eb8$sci),2),
  "Q3"=round(quantile(eb8$sci, probs = 0.75),2),
  "Maximum"=round(max(eb8$sci),2),
  "Std.deviation"=round(sd(eb8$sci),2)
  )

descriptive.sci8 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,science8-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.sci8 = cut(eb8$sci,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.sci8)*100/nrow(eb8)) %>% pander()
  
```

#### Density function

```{r,science8-density, echo=FALSE,fig.align='center'}

ggplot(eb8, aes(x=sci))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.9)+
  labs(x="score",
       y="probability",
       title="Standard 8",
       subtitle="Science:")+
  geom_vline(xintercept = mean(eb8$sci))

```

#### Best student(s)

```{r,science8-best, echo=FALSE,fig.align='center'}
eb8[eb8$sci == max(eb8$sci),] %>% pander()
```

### e. SSRE {.tabset .tabset-pill}

This is how Standard 8 pupils performed in SSRE.

#### Descriptive summaries

```{r,ssre8-1, echo=FALSE,fig.align='center'}
descriptive.ssre8 = data.frame(
  "Minimum"=round(min(eb8$ssre),2),
  "Q1"=round(quantile(eb8$ssre, probs = 0.25),2),
  "Mean"=round(mean(eb8$ssre),2),
  "Median"=round(median(eb8$ssre),2),
  "Q3"=round(quantile(eb8$ssre, probs = 0.75),2),
  "Maximum"=round(max(eb8$ssre),2),
  "Std.deviation"=round(sd(eb8$ssre),2)
  )

descriptive.ssre8 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,ssre8-2, echo=FALSE, fig.align='center'}
custom.breaks = c(0,10,20,30,40,50,60,70,80,90,100)
custom.labels = c("0-10","11-20","21-30","31-40","41-50",
                "51-60","61-70","71-80","81-90","91-100")

cut.ssre8 = cut(eb8$ssre,
                 breaks=custom.breaks, 
                 labels=custom.labels)

ceiling(table(cut.ssre8)*100/nrow(eb8)) %>% pander()
```

#### Density function

```{r,ssre8-density, echo=FALSE,fig.align='center'}

ggplot(eb8, aes(x=ssre))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.9)+
  labs(x="score",
       y="probability",
       title="Standard 8",
       subtitle="SSRE:")+
  geom_vline(xintercept = mean(eb8$ssre))

```

#### Best student(s)

```{r,ssre8-best, echo=FALSE,fig.align='center'}
eb8[eb8$ssre == max(eb8$ssre),] %>% pander()
```

### f. Total {.tabset .tabset-pill}

This is how Standard 8 pupils performed overally.

#### Descriptive summaries

```{r,total8-1, echo=FALSE,fig.align='center'}
descriptive.total8 = data.frame(
  "Minimum"=round(min(eb8$total),2),
  "Q1"=round(quantile(eb8$total, probs = 0.25),2),
  "Mean"=round(mean(eb8$total),2),
  "Median"=round(median(eb8$total),2),
  "Q3"=round(quantile(eb8$total, probs = 0.75),2),
  "Maximum"=round(max(eb8$total),2),
  "Std.deviation"=round(sd(eb8$total),2)
  )

descriptive.total8 %>% pander()
```

#### Score distribution

The table below represents the score distribution in percentage.

```{r,total8-2, echo=FALSE, fig.align='center'}
total.custom.breaks <-c(0,100,200,300,400,500) 
total.custom.labels <-c("0-100","101-200","201-300","301-400","401-500")

cut.total8 = cut(eb8$total,
                 breaks=total.custom.breaks, 
                 labels=total.custom.labels)

ceiling(table(cut.total8)*100/nrow(eb8)) %>% pander()
```

#### Density function

```{r,total8-density, echo=FALSE,fig.align='center'}

ggplot(eb8, aes(x=total))+
  geom_density(fill = "darkorange", col= "darkorange4", alpha = 0.9)+
  labs(x="score",
       y="probability",
       title="Standard 8",
       subtitle="Total:")+
  geom_vline(xintercept = mean(eb8$total))

```

#### Best student(s)

```{r,total8-best, echo=FALSE,fig.align='center'}
eb8[eb8$total == max(eb8$total),] %>% pander()
```

# Stream performance

Ebusamba Primary School has double streamed classes namely, *B* and *G*. This section looks to identify existing differences between subject performances between these two streams for the 3 classes. To do that, there is need to formulate research hypotheses and conduct tests that will not only help us identify differences but also signify them. First, we assume the following:

-   The pupils' exam scores from both streams and for all subjects are independent
-   The exam scores are normally distributed with respective means per stream and equal variances.

Given the above assumptions put in place, we shall use the ANOVA tests to investigate for performance differences. We set the level of significance to *5%*

## Research hypothesis {.tabset .tabset-pill}

> H0 = *There is no difference in subject performance between stream B and G*

### Standard 6 {.tabset .tabset-pill}

#### a. English

```{r,english6-aov, echo=FALSE}
aov.eng6=aov(data = eb6, formula = eng~strm)
aov.eng6 %>% summary() %>% pander()
```

Since p-value \> 0.05, we fail to reject the null hypothesis in favor of the initial claim that there is no difference in performance between stream B and G in English.

##### Boxplot

```{r, english6-box, echo=FALSE}
box.eng6=ggplot(data = eb6, aes(x=factor(strm), y=eng))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.5)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.eng6) %>% config(displayModeBar = FALSE)
```

#### b. Kiswahili

```{r,kiswahili6-aov, echo=FALSE}
aov.kis6=aov(data = eb6, formula = kis~strm)
aov.kis6 %>% summary() %>% pander()
```

The obtained p-value \< 0.05. In conclusion, we reject the research hypothesis and hence report a statistical significance in performance between stream B and G in Kiswahili.

##### Boxplot

```{r, kiswahili6-box, echo=FALSE}
box.kis6=ggplot(data = eb6, aes(x=factor(strm), y=kis))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.5)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.kis6) %>% config(displayModeBar = FALSE)
```

#### c. Mathematics

```{r,math6-aov, echo=FALSE}
aov.math6=aov(data = eb6, formula = math~strm)
aov.math6 %>% summary() %>% pander()
```

The obtained p-value \< 0.05. We reject the research hypothesis and report a statistical significance in performance between stream B and G in Mathematics.

##### Boxplot

```{r, math6-box, echo=FALSE}
box.math6=ggplot(data = eb6, aes(x=factor(strm), y=math))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.5)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.math6) %>% config(displayModeBar = FALSE)
```

#### d. Science

```{r,science6-aov, echo=FALSE}
aov.sci6=aov(data = eb6, formula = sci~strm)
aov.sci6 %>% summary() %>% pander()
```

The obtained p-value \< 0.05. We reject the research hypothesis and report a statistical significance in performance between stream B and G in Science.

##### Boxplot

```{r, science6-box, echo=FALSE}
box.sci6=ggplot(data = eb6, aes(x=factor(strm), y=sci))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.5)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.sci6) %>% config(displayModeBar = FALSE)
```

#### e. SSRE

```{r,ssre6-aov, echo=FALSE}
aov.ssre6=aov(data = eb6, formula = ssre~strm)
aov.ssre6 %>% summary() %>% pander()
```

The p-value \> 0.05. We fail to reject the null hypothesis. In conclusion, there is no difference in performance between stream B and G in SSRE.

##### Boxplot

```{r, ssre6-box, echo=FALSE}
box.ssre6=ggplot(data = eb6, aes(x=factor(strm), y=ssre))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.5)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.ssre6) %>% config(displayModeBar = FALSE)
```

#### f. Total

```{r,total6-aov, echo=FALSE}
aov.total6=aov(data = eb6, formula = total~strm)
aov.total6 %>% summary() %>% pander()
```

The obtained p-value \< 0.05. We reject the null hypothesis and report a statistical significance in the overall performance between stream B and G.

##### Boxplot

```{r, total6-box, echo=FALSE}
box.total6=ggplot(data = eb6, aes(x=factor(strm), y=total))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.5)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.total6) %>% config(displayModeBar = FALSE)
```

### Standard 7 {.tabset .tabset-pill}

#### a. English

```{r,eng7-aov, echo=FALSE}
aov.eng7=aov(data = eb7, formula = eng~strm)
aov.eng7 %>% summary() %>% pander()
```

Since p-value \< 0.05, we reject the null hypothesis in favor of the alternative hypothesis that there is a difference in performance between stream B and G in English.

##### Boxplot

```{r, english7-box, echo=FALSE}
box.eng7=ggplot(data = eb7, aes(x=factor(strm), y=eng))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.7)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.eng7) %>% config(displayModeBar = FALSE)
```

#### b. Kiswahili

```{r,kis7-aov, echo=FALSE}
aov.kis7=aov(data = eb7, formula = kis~strm)
aov.kis7 %>% summary() %>% pander()
```

Since p-value \> 0.05, we fail to reject the null hypothesis in favor of the initial claim that there is no difference in performance between stream B and G in Kiswahili.

##### Boxplot

```{r, kiswahili7-box, echo=FALSE}
box.kis7=ggplot(data = eb7, aes(x=factor(strm), y=kis))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.7)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.kis7) %>% config(displayModeBar = FALSE)
```

#### c. Mathematics

```{r,math7-aov, echo=FALSE}
aov.math7=aov(data = eb7, formula = math~strm)
aov.math7 %>% summary() %>% pander()
```

The obtained p-value \> 0.05. We fail to reject the research hypothesis. We conclude a statistical insignificance in the subject performance of Mathematics between stream B and G.

##### Boxplot

```{r, math7-box, echo=FALSE}
box.math7=ggplot(data = eb7, aes(x=factor(strm), y=math))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.7)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.math7) %>% config(displayModeBar = FALSE)
```

#### d. Science

```{r, science7-aov, echo=FALSE}
aov.sci7=aov(data = eb7, formula = sci~strm)
aov.sci7 %>% summary() %>% pander()
```

The obtained p-value \> 0.05. We fail to reject the research hypothesis. We conclude a statistical insignificance in the subject performance of Science between stream B and G.

##### Boxplot

```{r, science7-box, echo=FALSE}
box.sci7=ggplot(data = eb7, aes(x=factor(strm), y=sci))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.7)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.sci7) %>% config(displayModeBar = FALSE)
```

#### e. SSRE

```{r, ssre7-aov, echo=FALSE}
aov.ssre7=aov(data = eb7, formula = ssre~strm)
aov.ssre7 %>% summary() %>% pander()
```

The obtained p-value \> 0.05. We fail to reject the research hypothesis. We conclude a statistical insignificance in the subject performance of SSRE between stream B and G.

##### Boxplot

```{r, ssre7-box, echo=FALSE}
box.ssre7=ggplot(data = eb7, aes(x=factor(strm), y=ssre))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.7)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.ssre7) %>% config(displayModeBar = FALSE)
```

#### f. Total

```{r, total7-aov, echo=FALSE}
aov.total7=aov(data = eb7, formula = total~strm)
aov.total7 %>% summary() %>% pander()
```

The obtained p-value \> 0.05. We fail to reject the research hypothesis. In conclusion, there is no difference in the overall performance between the 2 streams.

##### Boxplot

```{r, total7-box, echo=FALSE}
box.total7=ggplot(data = eb7, aes(x=factor(strm), y=total))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.7)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.total7) %>% config(displayModeBar = FALSE)
```

### Standard 8 {.tabset .tabset-pill}

#### a. English

```{r, eng8-aov, echo=FALSE}
aov.eng8=aov(data = eb8, formula = eng~strm)
aov.eng8 %>% summary() %>% pander()
```

The obtained p-value \> 0.05. We fail to reject the null hypothesis. In conclusion, there is no difference in the subject performance of English between stream B and G.

##### Boxplot

```{r, eng8-box, echo=FALSE}
box.eng8=ggplot(data = eb8, aes(x=factor(strm), y=eng))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.9)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.eng8) %>% config(displayModeBar = FALSE)
```

#### b. Kiswahili

```{r, kis8-aov, echo=FALSE}
aov.kis8=aov(data = eb8, formula = kis~strm)
aov.kis8 %>% summary() %>% pander()
```

The obtained p-value \> 0.05. We fail to reject the research hypothesis. In conclusion, there is no difference in the Kiswahili subject performance between the 2 streams.

##### Boxplot

```{r, kis8-box, echo=FALSE}
box.kis8=ggplot(data = eb8, aes(x=factor(strm), y=kis))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.9)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.kis8) %>% config(displayModeBar = FALSE)
```

#### c. Mathematics

```{r, math8-aov, echo=FALSE}
aov.math8=aov(data = eb8, formula = math~strm)
aov.math8 %>% summary() %>% pander()
```

The obtained p-value \> 0.05. We fail to reject the research hypothesis.There is no difference in the performance of Mathematics between the stream B and G.

##### Boxplot

```{r, math8-box, echo=FALSE}
box.math8=ggplot(data = eb8, aes(x=factor(strm), y=math))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.9)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.math8) %>% config(displayModeBar = FALSE)
```

#### d. Science

```{r, science8-aov, echo=FALSE}
aov.sci8=aov(data = eb8, formula = sci~strm)
aov.sci8 %>% summary() %>% pander()
```

The obtained p-value \> 0.05. We fail to reject the research hypothesis. In conclusion, there is no difference in the Science subject performance between the 2 streams.

##### Boxplot

```{r, science8-box, echo=FALSE}
box.sci8=ggplot(data = eb8, aes(x=factor(strm), y=sci))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.9)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.sci8) %>% config(displayModeBar = FALSE)
```

#### e. SSRE

```{r, ssre8-aov, echo=FALSE}
aov.ssre8=aov(data = eb8, formula = ssre~strm)
aov.ssre8 %>% summary() %>% pander()
```

The obtained p-value \> 0.05. We fail to reject the research hypothesis. In conclusion, there is no difference in the SSRE subject performance between the 2 streams.

##### Boxplot

```{r, ssre8-box, echo=FALSE}
box.ssre8=ggplot(data = eb8, aes(x=factor(strm), y=ssre))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.9)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.ssre8) %>% config(displayModeBar = FALSE)
```

#### f. Total

```{r, total8-aov, echo=FALSE}
aov.total8=aov(data = eb8, formula = total~strm)
aov.total8 %>% summary() %>% pander()
```

The obtained p-value \> 0.05. We fail to reject the research hypothesis. In conclusion, there is no difference in the overall performance between stream B and G.

##### Boxplot

```{r, total8-box, echo=FALSE}
box.total8=ggplot(data = eb8, aes(x=factor(strm), y=total))+
  geom_boxplot(fill = "darkorange", col= "darkorange4", alpha = 0.9)+
  stat_summary(fun.y = "mean", geom = "point", shape = 23, size = 2.5, fill = "darkorange", col = "darkorange4")+
  labs(x = "streams", y = "score")

ggplotly(box.total8) %>% config(displayModeBar = FALSE)
```

# Findings {.tabset .tabset-pill}

## Standard 6 {.tabset .tabset-pill}

-   The candidates recorded the highest variation in Kiswahili and Science (sd = 14.27)

-   Mathematics had the least variation (sd = 8.88)

-   Mathematics was the least performed subject (mean = 27.26)

-   Kiswahili is the best performed subject (mean = 46.27)

-   A majority of the standard 6 candidates scored an overall of 101-200 marks (71%)

-   Subject performance between the two streams was significantly different in Kiswahili, Mathematics and Science.

-   Overall performance between the two streams was also significant.

-   All the best students per subject emerge from the B stream.

## Standard 7 {.tabset .tabset-pill}

-   The candidates recorded the highest variation in Science (sd = 13.25)

-   SSRE had the least variation (sd = 10.83)

-   Mathematics was the least performed subject (mean = 34.30)

-   Science was the best performed subject (mean = 45.41)

-   Numerous standard 7 candidates scored an overall of 101-200 marks (60%)

-   Subject performance between the two streams was insignificantly different for all the subjects.

## Standard 8 {.tabset .tabset-pill}

-   The candidates recorded the highest variation in Science (sd = 12.51)

-   The least variation was recorded in Kiswahili (sd = 9.71)

-   Mathematics was the least performed subject (mean = 32.10)

-   English was the best performed subject (mean = 46.54)

-   Multiple standard 8 candidates scored an overall of 201-300 marks (54%)

-   Subject performance between the two streams was insignificantly different for all the subjects.

# Recommendation

-   Science teachers should come up with common strategies for disseminating knowledge to the learners in order to kurb variation in the subject performance across the three classes.

-   Introduce Mathematical discussion groups amongst the candidates in each stream.

-   Reward top performers to enhance healthy academic competition with other learners.
