---
title: "Premier League 2023/2024: Manchester City"
output: 
  flexdashboard::flex_dashboard:
    theme: 
      bg: "#FFFFFF"
      fg: "#1C2C5B" 
      primary: "#6CABDD"
    orientation: columns
    vertical_layout: scroll
    social: [twitter, linkedin]
    favicon: IMG_20240907_141449.jpg
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(message=FALSE, warning=FALSE, tidy.opts = list(width.cutoff=60))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(shiny))
suppressPackageStartupMessages(library(flexdashboard))
suppressPackageStartupMessages(library(pander))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(ggvis))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(ggplotlyExtra))

epl=data.frame(read.csv("matches.csv", header=TRUE))
mci<-epl[epl$Team=="ManchesterCity",]
```


Introduction
=============================================================

Column
---------------------------------------------------------------

Manchester City is a professional football club based in Manchester, England. The team took part in the 2023/24 Premier League season and won their fourth successive title under manager Pep Guardiola. The Cityzens took their tally of Premier League trophies to 7. They also booked a spot in next season's revamped UEFA Champions League.This dashboard looks at the club's on-the-pitch performance in the English top flight division.

### Pep Guardiola

```{r}
include_graphics(
"IMG_20240907_141531.jpg")
```


Column
---------------------------------------------------------------
### 

```{r}
include_graphics(
"IMG_20240907_141512.jpg")
```


### 

```{r}
include_graphics(
"wp11075864-etihad-stadium-2022-wallpapers.jpg")
```

Players {data-orientation="rows"}
===========================================================

Row
------------------------------------------------------------
```{r}
## PLAYERS' DATASET
sheets<-excel_sheets("Soccer-Stats-Premier-League-2023-2024_Final_20240610.xlsx")
players<-data.frame(read_excel("Soccer-Stats-Premier-League-2023-2024_Final_20240610.xlsx",
                               sheet = sheets[4]))

# dim(players) # 746 rows by 27 columns
# Checking for duplicates  
# players[duplicated(players), ]  # No duplicates found
# Checking for missing values
# players[!complete.cases(players), ] # 572 cases with NAs  
#players<-players[complete.cases(players), ]# 487 complete cases
players1<-players[,c(2,7,9,10,12,20,23,24,25,26)]

# DATA TYPE CONVERTION
players1$Name<-factor(players1$Name)
players1$Team<-factor(players1$Team)
players1$Position<-factor(players1$Position)
```

### Registered players

```{r}
players %>% 
  filter(Team == "Manchester City") %>% 
  count() %>% 
  valueBox(color="#FFC659")
```

### Goalkeepers

```{r}
players %>% 
  filter(Team == "Manchester City",
         Position == "Goalkeeper") %>% 
  count() %>% 
  valueBox(color="#FFC659")
```

### Defenders

```{r}
players %>% 
  filter(Team == "Manchester City",
         Position == "Defender") %>% 
  count() %>% 
  valueBox(color = "#FFC659")
```

### Midfielders

```{r}
players %>% 
  filter(Team == "Manchester City",
         Position == "Midfielder") %>% 
  count() %>% 
  valueBox(color="#FFC659")
```

### Forwards

```{r}
players %>% 
  filter(Team == "Manchester City",
         Position == "Forward") %>% 
  count() %>% 
  valueBox(color="#FFC659")
```

### Average age (years)

```{r}
players %>% 
  filter(Team == "Manchester City") %>%
  summarise(round(mean(Age),0)) %>% 
  valueBox(color="#FFC659")
```

Row {.tabset .tabset-fade}
----------------------------------------------------------------

### Goalkeepers

```{r}
players %>% 
  filter(Team == "Manchester City",
         Position == "Goalkeeper") %>% 
  select(Citizenship, Name, Jersey) %>% 
  pander()
```

### Defenders

```{r}
players %>% 
  filter(Team == "Manchester City",
         Position == "Defender") %>% 
  select(Citizenship, Name, Jersey) %>% 
  pander()
```

### Midfielders

```{r}
players %>% 
  filter(Team == "Manchester City",
         Position == "Midfielder") %>% 
  select(Citizenship, Name, Jersey) %>% 
  pander()
```

### Forwards

```{r}
players %>% 
  filter(Team == "Manchester City",
         Position == "Forward") %>% 
  select(Citizenship, Name, Jersey) %>% 
  pander()
```


Outcomes
=========================================================

Column
-------------------------------------------------------
### Home


```{r}
cat("Wins","[%]")
#mci %>% 
  #filter(Venue == "Home" & Result=="W") %>% 
  #summarise(win.pct.h=n()*100/19) %>% 
  gauge(73.68, min=0,max=100,symbol="%",
        gaugeSectors(success = c(80,100),
                     warning =c(40,79),
                     danger = c(0,39)))

```

### 

```{r}
cat("Draws","[%]")
#mci %>% 
  #filter(Venue == "Home" & Result=="D") %>% 
  #summarise(draw.pct.h=n()*100/19) %>% 
  gauge(26.31, min=0,max=100,symbol="%",
        gaugeSectors(success = c(80,100),
                     warning =c(40,79),
                     danger = c(0,39)))

```

### 

```{r}
cat("Losses","[%]")
#mci %>% 
  #filter(Venue == "Home" & Result=="L") %>% 
  #summarise(loss.pct.h=n()*100/19) %>% 
  gauge(0, min=0,max=100,symbol="%",
        gaugeSectors(success = c(0,39),
                     warning =c(40,79),
                     danger = c(80,100)))

```

Column
-------------------------------------------------------
### Away


```{r}
cat("Wins","[%]")
#mci %>% 
  #filter(Venue == "Away" & Result=="W") %>% 
  #summarise(win.pct.a=n()*100/19) %>% 
  gauge(73.68, min=0,max=100,symbol="%",
        gaugeSectors(success = c(80,100),
                     warning =c(40,79),
                     danger = c(0,39)))

```

### 

```{r}
cat("Draws","[%]")
#mci %>% 
  #filter(Venue == "Away" & Result=="D") %>% 
  #summarise(draw.pct.a=n()*100/19) %>% 
  gauge(10.53, min=0,max=100,symbol="%")

```

### 

```{r}
cat("Losses","[%]")
#mci %>% 
  #filter(Venue == "Away" & Result=="L") %>% 
  #summarise(loss.pct.a=n()*100/19) %>% 
  gauge(15.79, min=0,max=100,symbol="%",
        gaugeSectors(success = c(0,39),
                     warning =c(40,79),
                     danger = c(80,100)))

```

Awards
===========================================================

Column
------------------------------------------------------------
### Goals scored

```{r, fig.align="top", out.width="10%"}
include_graphics(
"IMG_20240908_223341.jpg")
```

### 
```{r}
players1 %>%
  filter(Team == "Manchester City",
         totalGoals >= 1) %>% 
  select(Name, totalGoals) %>%
  arrange(desc(totalGoals)) %>% 
  pander()
  
```

Column
------------------------------------------------------------
### Goal Assists

```{r, fig.align="top", out.width="10%"}
include_graphics(
"IMG_20240907_140408.jpg")
```

###
```{r}
players1 %>%
  filter(Team == "Manchester City",
         goalAssists >= 1) %>% 
  select(Name, goalAssists) %>%
  arrange(desc(goalAssists)) %>% 
  pander()
  
```


The Etihad
===========================================================

Column 
-----------------------------------------------------------
### 

```{r}
include_graphics(
"IMG_20240907_141933.jpg")
```

### 

```{r, fig.align="bottom", out.width="20%"}
include_graphics(
"wp10675216-manchester-city-stadium-wallpapers.jpg")
```


Column{.tabset .tabset-fade}
----------------------------------------------------------------------

### Notes

The Etihad played host to 19 Premier League home games over the 2023/24 campaign. The average match attendance stands at **55,012**.

The highest attendance was recorded in **May 2024** against **West Ham** (full attendance) and the lowest at **52,143** against **Burnley** in **January 2024**.



### Home matches

```{r}
mci.home<-
  mci[mci$Venue == "Home",]%>% 
  select(c(Date,Round,Opponent,Attendance)) %>% 
  arrange(desc(Attendance)) %>%
  mutate(percent =   round(Attendance*100/(55097),2))%>% 
  datatable(filter="none",rownames=FALSE ,options=list(searching=TRUE, pagelength=5)); mci.home  
```

### Distribution

```{r}
mci.at.home<-mci[mci$Venue=="Home", ]
mci.attendance<-ggplot(mci.at.home, aes(x=Date,y=Attendance))+
  geom_point(shape=20, size=4, color="#6CABDD")+
  geom_hline(aes(yintercept=round(mean(Attendance),0)))+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(x="Time",
       y="Attendance",
       title="Premier League\nhome attendance: The Etihad 2023/24")+
  theme_pander()

plotly::ggplotly(p=mci.attendance)
```


Goals scored {data-navmenu="Offensive record"}
======================================================================

```{r , fig.width=14, fig.height=5}
offensive1<-epl %>% 
  group_by(Team) %>% 
  summarise(goals_scored = sum(GF)) %>% 
  arrange(desc(goals_scored)) %>% 
  head(10)  

GS<-ggplot(offensive1,aes(x=reorder(Team, goals_scored), y=goals_scored))+
  geom_bar(stat='identity',fill = "#6CABDD", alpha=0.85)+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
 labs(x="Teams",
      y="Goals scored",
      title="Goals scored:\nPremier League 2023/2024",
      subtitle = "Top 10")+
  theme_pander()

ggplotly(GS)
```

Column
----------------------------------------------------------------------

### Position

```{r,fig.width=4}
epl %>% 
  group_by(Team) %>% 
  summarise(total.goals = sum(GF)) %>%
  arrange(desc(total.goals)) %>% 
  mutate(position = min_rank(-total.goals)) %>% 
  filter(Team == "ManchesterCity") %>% 
  select(position) %>% 
  valueBox(color="#FFC659")
```

### Goals scored

```{r,fig.width=4}
valueBox(value=sum(mci$GF),color="#FFC659")
```


Column
-------------------------------------------------------------------------

### Home Goals

```{r, fig.width=4}

valueBox(value = sum(mci.at.home$GF),color="#FFC659")
```

### Away Goals

```{r, fig.width=4}
valueBox(value=sum(mci$GF)-sum(mci.at.home$GF),color="#FFC659")
```



# Shots {data-navmenu="Offensive record"}

```{r ,fig.width=14, fig.height=5}
offensive2<-epl %>% 
  group_by(Team) %>%
  summarise(Shot_accuracy = sum(SoT)*100/sum(Sh)) %>% 
  arrange(desc(Shot_accuracy)) %>% 
  head(10) 

SOT<-ggplot(offensive2 ,aes(x=reorder(Team, Shot_accuracy), y=Shot_accuracy))+
  geom_bar(stat='identity',fill = "#6CABDD",alpha=0.85)+
  theme_pander()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(x="Teams",
       y="Percentage",
       title="Shot accuracy [%]:\nPremier League 2023/2024",
       subtitle = "Top 10")

ggplotly(SOT)
```

Column
----------------------------------------------------------------------

### Position

```{r,fig.width=4}
epl %>% 
  group_by(Team) %>% 
  summarise(shot.acc = sum(SoT)*100/sum(Sh)) %>%
  arrange(desc(shot.acc)) %>% 
  mutate(position = min_rank(-shot.acc)) %>% 
  filter(Team == "ManchesterCity") %>% 
  select(position) %>%
valueBox(color="#FFC659")
```

### Shot accuracy [%]

```{r,fig.width=4}
valueBox(value=round(sum(mci$SoT)*100/sum(mci$Sh), 2)
, color="#FFC659")
```


Possession {data-navmenu="Offensive record"}
=====================================================================

```{r ,fig.width=14, fig.height=5}
offensive3<-epl %>% 
  group_by(Team) %>%
  summarise(avg.poss = mean(Poss)) %>% 
  arrange(desc(avg.poss)) %>% 
  head(10)  

POSS<-ggplot(offensive3, aes(x=reorder(Team, avg.poss), y=avg.poss))+
  geom_bar(stat='identity',fill = "#6CABDD",alpha=0.85)+
  theme_pander()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(x="Teams",
       y="Percentage",
       title="Average ball possession [%]:\nPremier League 2023/2024",
       subtitle = "Top 10")

ggplotly(POSS)
```

Column
----------------------------------------------------------------------

### Position

```{r,fig.width=4}
epl %>% 
  group_by(Team) %>% 
  summarise(avg.poss = mean(Poss)) %>%
  arrange(desc(avg.poss)) %>% 
  mutate(position = min_rank(-avg.poss)) %>% 
  filter(Team == "ManchesterCity") %>% 
  select(position) %>% 
  valueBox(color="#FFC659")
```

### Average possesion [%]

```{r,fig.width=4}
valueBox(value=round(mean(mci$Poss), 2)
, color="#FFC659")
```


Goals conceded {data-navmenu="Defensive record"}
===========================================================

```{r ,fig.width=14, fig.height=5}
defensive1<-epl %>% 
  group_by(Team) %>%
  summarise(gc = sum(GA)) %>% 
  arrange(gc) %>% 
  head(10) 
CONC<-ggplot(defensive1, aes(x=reorder(Team, gc), y=gc))+
  geom_bar(stat='identity', fill = "#6CABDD",alpha=0.85)+
  theme_pander()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(x="Teams",
       y="goals",
       title="Goals conceded:\nPremier League 2023/2024",
       subtitle = "Top 10:Least conceded goals")

ggplotly(CONC)
```

Column
----------------------------------------------------------------------

### Position

```{r,fig.width=4}
valueBox(value=2,  color="#FFC659")
```

### Goals conceded

```{r,fig.width=4}
valueBox(value=sum(mci$GA), color="#FFC659")
```

Column
----------------------------------------------------------------------

### At Home

```{r, fig.width=4}

valueBox(value = sum(mci.at.home$GA), 
         color="#FFC659")
```

### Away

```{r, fig.width=4}
valueBox(value=sum(mci$GA)-sum(mci.at.home$GA), 
         color="#FFC659")
```


Clean sheets {data-navmenu="Defensive record"}
========================================================

```{r,fig.width=14, fig.height=5}
defensive2<-epl[epl$GA == 0, ] %>% 
  group_by(Team) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(10)
  
CS<-ggplot(defensive2,aes(x=reorder(Team, n), y=n))+
  geom_bar(stat='identity', fill = "#6CABDD",alpha=0.85)+
  theme_pander()+
  theme(axis.text.x = element_text(angle=45, hjust=1))+
  labs(x="Teams",
       y="Frequency",
       title="Clean sheets:\nPremier League 2023/2024") 

ggplotly(CS)
```

Column
----------------------------------------------------------------------

### Position

```{r}
epl[epl$GA == 0, ] %>% 
  group_by(Team) %>% 
  summarise(c.s = n()) %>%
  mutate(position = min_rank(-c.s)) %>%
  arrange(desc(c.s)) %>% 
  filter(Team == "ManchesterCity") %>% 
  select(position) %>%
valueBox(color="#FFC659")
```

### Clean sheets

```{r}
mci[mci$GA==0,] %>% 
  count() %>% 
  valueBox(color="#FFC659")
```

Column
----------------------------------------------------------------------

### At home

```{r}
mci[mci$Venue == "Home",] %>%
  filter(GA == 0) %>% 
  nrow() %>%
valueBox(color="#FFC659")
```

### Away

```{r}
# 13-6=7
mci[mci$Venue == "Away",] %>%
  filter(GA == 0) %>% 
  nrow() %>%
valueBox(color="#FFC659")
```


References
============================================================

- [www.kaggle.com/datasets/mertbayraktar/english-premier-league-matches-20232024-season](www.kaggle.com/datasets/mertbayraktar/english-premier-league-matches-20232024-season)

- [https://pkgs.rstudio.com/flexdashboard/](https://pkgs.rstudio.com/flexdashboard/)

- [https://www.excel4soccer.com/
](https://www.excel4soccer.com/
)

